\LARGE \bf Geometric Adaptive Control of Attitude Dynamics on $\SO$\\ with State Inequality Constraints

### 2015-10-23 ACC16.2.tex
fixed another error to reduce page length. now this is final. for reals
corrected typo

Compared to the previous work on constrained attitude control, we present a geometrically exact control system without parameterizations.
In addition, we incorporate state inequality constraints for the first time on \( \SO \).
The presented control system is computed in real-time and offers significant computational advantages over previous iterative methods. 
In addition, the riguous mathematical proof guarantees stability.
\section{Numerical Examples}
We demonstrate the performance of the proposed control system via numerical simulation.
The inertia tensor of a rigid body is given as
\begin{gather*}
	J = \begin{bmatrix}
	\num{5.57e-3} & \num{6.17e-5} & \num{-2.50e-5} \\
	\num{6.17e-5} & \num{5.57e-3} & \num{1.00e-5} \\
	\num{-2.50e-5} & \num{1.00e-5} & \num{1.05e-2}
	\end{bmatrix} \si{\kilo\gram\meter\squared} .
\end{gather*} 
The control system parameters are chosen as
\begin{gather*}
	G = \text{diag} [0.9,1.1,1.0], \quad k_R = 0.4 , \quad	k_\Omega = 0.296 ,\\
	c = 1.0 , \quad k_\Delta = 0.5 , \quad \alpha = 15 .
\end{gather*}
A body fixed sensor is defined as \(r = [1,0,0]\), while multiple inequality constraints are defined in~\Cref{tab:constraints}.
The simulation parameters are chosen to be similar to those found in~\cite{lee2011b}, howver we increase the size of the constraint regions to create a more challenging scenario for the control system.

@@

%To show~\cref{item:prop_crit}, we determine the critical points of~\cref{eqn:eRA}.
%The attitude error between \( R \) and \( R_d \) is defined by \( Q \in \SO \) as
%\begin{align*}
%	Q = R_d^T R = 
%	\begin{bmatrix}
%		q_{11} & q_{12} & q_{13} \\
%		q_{21} & q_{22} & q_{23} \\
%		q_{31} & q_{32} & q_{33}
%	\end{bmatrix} .
%\end{align*}
%Equating \( e_{R_{A}} = 0 \) yields the following equivalent relations
%\begin{subequations}
%\begin{align}
%	g_3 q_{32} &= g_2 q_{23} , \\
%	g_1 q_{13} &= g_3 q_{31} , \\
%	g_2 q_{21} &= g_1 q_{12} .
%\end{align}
%\end{subequations}
%These relations allow one to redefine the attitude error \( Q \) in terms of the independent upper diagonal elements.
%Noting that \( Q \in \SO \) is an orthonormal matrix, with unit length row and columns gives
%\begin{subequations}
%\begin{align}
%	\parenth{1 - \frac{g_1^2}{g_2^2}} q_{12}^2 + \parenth{1 - \frac{g_1^2}{g_3^2}}q_{13}^2 &= 0 , \\
%	\parenth{1 - \frac{g_1^2}{g_2^2}} q_{12}^2 - \parenth{1 - \frac{g_2^2}{g_3^2}} q_{23}^2 &= 0 , \\
%	\parenth{1-\frac{g_1^2}{g_3^2}} q_{13}^2 + \parenth{1 - \frac{g_2^2}{g_3^2}} q_{23}^2 &= 0 .
%\end{align}
%\end{subequations}
%As a result, all the off diagonal terms of \( Q = R_d^T R \) are zero, namely \( q_{12}= q_{13} = q_{23} = 0 \).
%Since \( Q \in \SO \), the diagonal elements \( q_{11}, q_{22}, \text{ and } q_{33} \) are either \( +1 \) or \(-1\).
The critical points of \( e_{R_A} \) are derived in~\cite{bullo2004}.
There are four critical points of \( e_{R_A} \), the desired attitude \( R_d \) as well as rotations about each body fixed axis by \ang{180}.
The repulsive error vector \( e_{R_{B}} \) is zero only when the numerator \( \parenth{R^T v}^\wedge r = 0 \). 
This condition only occurs if the desired attitude results in the body fixed vector \( r \) becoming aligned with \(R^T v \) while simultaneously satisfying \(\braces{R_d} \cup \braces{R_d \exp(\pi \hat{s}} \) for \( s \in \braces{e_1, e_2, e_3} \).
Since we assume the system will not operate in violation of the constraints, the addition of the barrier function does not add additional critical points to the control system.
The desired equilibrium is \( e_R = 0 \) and \( A = 0\).
The proof of \( A(R) \) given by~\cref{item:prop_era_upbound} in~\Cref{prop:config_error} is available in~\cite{LeeITCST13}.
\subsection{Proof of~\Cref{prop:error_dyn}}\label{proof:error_dyn}
From the kinematics~\cref{eqn:Rdot} and noting that \( \dot{R}_d = 0 \) the time derivative of \( R_d^T R \) is given as
\begin{gather*}
	\diff{}{t} \parenth{R_d^T R} = R_d^T R \hat{e}_\Omega .
\end{gather*}
Applying this to the time derivative of~\cref{eqn:A} gives
\begin{gather*}
	\diff{}{t} (A) = -\frac{1}{2} \tr{G R_d^T R \hat{e}_\Omega} .
\end{gather*}
Applying~\cref{eqn:hat1} into this shows~\cref{eqn:A_dot}.
Next, the time derivative of the repulsive error function is given by
\begin{gather*}
	\diff{}{t} (B) = \frac{r^T \parenth{\hat{\Omega} R^T} v}{\alpha \parenth{r^T R^T v - \cos \theta}} .
\end{gather*}
Using the scalar triple product, given by~\cref{eqn:STP}, one can reduce this to~\cref{eqn:B_dot}.
The time derivative of the attractive attitude error vector, \( e_{R_A} \), is given by
\begin{gather*}
	\diff{}{t} ( e_{R_A}) = \frac{1}{2} \parenth{\hat{e}_\Omega R^T R_d G + (R^T R_d G)^T \hat{e}_\Omega}^\vee .
\end{gather*}
Using the hat map property given in~\cref{eqn:xAAx} this is further reduced to~\cref{eqn:eRA_dot,eqn:E}.

>>>>

Compared to the previous work on constrained attitude control, we present a geometrically exact control system without parameterizations.
In addition, we incorporate state inequality constraints for the first time on \( \SO \).
The presented control system is computed in real-time and offers significant computational advantages over previous iterative methods. 
In addition, the riguous mathematical proof guarantees stability.
\section{Numerical Examples}
We demonstrate the performance of the proposed control system via numerical simulation.
The inertia tensor of a rigid body is given as
\begin{gather*}
	J = \begin{bmatrix}
	\num{5.57e-3} & \num{6.17e-5} & \num{-2.50e-5} \\
	\num{6.17e-5} & \num{5.57e-3} & \num{1.00e-5} \\
	\num{-2.50e-5} & \num{1.00e-5} & \num{1.05e-2}
	\end{bmatrix} \si{\kilo\gram\meter\squared} .
\end{gather*} 
The control system parameters are chosen as
\begin{gather*}
	G = \text{diag} [0.9,1.1,1.0], \quad k_R = 0.4 , \quad	k_\Omega = 0.296 ,\\
	c = 1.0 , \quad k_\Delta = 0.5 , \quad \alpha = 15 .
\end{gather*}
A body fixed sensor is defined as \(r = [1,0,0]\), while multiple inequality constraints are defined in~\Cref{tab:constraints}.
The simulation parameters are chosen to be similar to those found in~\cite{lee2011b}, however we increase the size of the constraint regions to create a more challenging scenario for the control system.

@@

%To show~\cref{item:prop_crit}, we determine the critical points of~\cref{eqn:eRA}.
%The attitude error between \( R \) and \( R_d \) is defined by \( Q \in \SO \) as
%\begin{align*}
%	Q = R_d^T R = 
%	\begin{bmatrix}
%		q_{11} & q_{12} & q_{13} \\
%		q_{21} & q_{22} & q_{23} \\
%		q_{31} & q_{32} & q_{33}
%	\end{bmatrix} .
%\end{align*}
%Equating \( e_{R_{A}} = 0 \) yields the following equivalent relations
%\begin{subequations}
%\begin{align}
%	g_3 q_{32} &= g_2 q_{23} , \\
%	g_1 q_{13} &= g_3 q_{31} , \\
%	g_2 q_{21} &= g_1 q_{12} .
%\end{align}
%\end{subequations}
%These relations allow one to redefine the attitude error \( Q \) in terms of the independent upper diagonal elements.
%Noting that \( Q \in \SO \) is an orthonormal matrix, with unit length row and columns gives
%\begin{subequations}
%\begin{align}
%	\parenth{1 - \frac{g_1^2}{g_2^2}} q_{12}^2 + \parenth{1 - \frac{g_1^2}{g_3^2}}q_{13}^2 &= 0 , \\
%	\parenth{1 - \frac{g_1^2}{g_2^2}} q_{12}^2 - \parenth{1 - \frac{g_2^2}{g_3^2}} q_{23}^2 &= 0 , \\
%	\parenth{1-\frac{g_1^2}{g_3^2}} q_{13}^2 + \parenth{1 - \frac{g_2^2}{g_3^2}} q_{23}^2 &= 0 .
%\end{align}
%\end{subequations}
%As a result, all the off diagonal terms of \( Q = R_d^T R \) are zero, namely \( q_{12}= q_{13} = q_{23} = 0 \).
%Since \( Q \in \SO \), the diagonal elements \( q_{11}, q_{22}, \text{ and } q_{33} \) are either \( +1 \) or \(-1\).
The critical points of \( e_{R_A} \) are derived in~\cite{bullo2004}.
There are four critical points of \( e_{R_A} \), the desired attitude \( R_d \) as well as rotations about each body fixed axis by \ang{180}.
The repulsive error vector \( e_{R_{B}} \) is zero only when the numerator \( \parenth{R^T v}^\wedge r = 0 \). 
This condition only occurs if the desired attitude results in the body fixed vector \( r \) becoming aligned with \(R^T v \) while simultaneously satisfying \(\braces{R_d} \cup \braces{R_d \exp(\pi \hat{s}} \) for \( s \in \braces{e_1, e_2, e_3} \).
Since we assume the system will not operate in violation of the constraints, the addition of the barrier function does not add additional critical points to the control system.
The desired equilibrium is \( e_R = 0 \) and \( A = 0\).
The proof of \( \norm{e_{R_A}} \) given by~\cref{item:prop_era_upbound} is available in~\cite{LeeITCST13}.
\subsection{Proof of~\Cref{prop:error_dyn}}\label{proof:error_dyn}
From the kinematics~\cref{eqn:Rdot} and noting that \( \dot{R}_d = 0 \) the time derivative of \( R_d^T R \) is given as
\begin{gather*}
	\diff{}{t} \parenth{R_d^T R} = R_d^T R \hat{e}_\Omega .
\end{gather*}
Applying this to the time derivative of~\cref{eqn:A} gives
\begin{gather*}
	\diff{}{t} (A) = -\frac{1}{2} \tr{G R_d^T R \hat{e}_\Omega} .
\end{gather*}
Applying~\cref{eqn:hat1} into this shows~\cref{eqn:A_dot}.
Next, the time derivative of the repulsive error function is given by
\begin{gather*}
	\diff{}{t} (B) = \frac{r^T \parenth{\hat{\Omega} R^T} v}{\alpha \parenth{r^T R^T v - \cos \theta}} .
\end{gather*}
Using the scalar triple product, given by~\cref{eqn:STP}, one can reduce this to~\cref{eqn:B_dot}.
The time derivative of the attractive attitude error vector, \( e_{R_A} \), is given by
\begin{gather*}
	\diff{}{t} ( e_{R_A}) = \frac{1}{2} \parenth{\hat{e}_\Omega R^T R_d G + (R^T R_d G)^T \hat{e}_\Omega}^\vee .
\end{gather*}
Using the hat map property given in~\cref{eqn:xAAx} this is further reduced to~\cref{eqn:eRA_dot,eqn:E}.
